\thesistitle

### 2019-06-12 sections/research.tex
Literal/variable variability

\subsection{Type 1R clones} \label{chap:type1rclones}
We propose an alternative definition of type 1 clones. This definition requires cloned fragments to be not just textually equal, but also functionally equal. Although requiring fragements to be functionally equal, type 1R clones do not allow for change in implementation (like type 4 clones). We check functional equality of two fragments by validating the equality of the fully qualityfied identifier for referenced types, methods and variables. Type 1R clones are always a subset of type 1 clones.

@@

Because of this, for type 1R clones, we compare the fully qualified method signature for all method references. A fully qualified method signature consists of the fully qualified name of the method, the fully qualified type of the method plus the fully qualified type of each of its arguments. For instance, an \texttt{eat} method could become \texttt{java.lang.Boolean com.simonbaars.fruitgame.Apple.eat(java.util.List<com.simonbaars.fruitgame.tools.Tool>)}.

\subsubsection{Variables}
In typed programming languages, each variable declaration should declare a name and a type. When we reference a variable, we only use its name. If, in different code fragments, we use variables with the same name but different types, the code can be functionally unequal but still textually equal. As an example, see the code in figure \ref{lst:type1variables}.

\begin{figure}[H]
  \includegraphics[width=1\columnwidth]{img/type2}
  \caption{Variables with different types but the same name.}
  \label{fig:type1variables}
\end{figure}

The body of both methods in figure \ref{lst:type2variables} is equal. However, their functionality is not. The first method adds two numbers together and the other concatenates an integer to a String.

@@

\section{Type 2R clones}
Type 2R clones are modelled after type 2 clones, which allow any change in identifiers, literals, types, layout, and comments. For refactoring purposes, this definition is unsuitable; if we allow any change in identifiers, literals, and types, we cannot distinguish between different variables, different types and different method calls anymore. This could render two methods that have an entirely different functionality as clones (as shown in figure \ref{fig:type2} previously). Refactoring such clones can be harmful instead of helpful.

We tackle these problems with type 2R clones to be able to detect such clones that can and should be refactored. Type 1R clones are a subset of type 2R clones. All rules that apply to type 1R clones also apply to type 2R clones. Additionally, type 2R clones allow variability in literals, variables and method calls. This variability however is constrained by a threshold. Type 2R do not allow any variability in types, as opposed to type 2 clones which do allow variability in types.

@@

\noindent \textbf{Literal variability}\\
We allow only variability in the value of literals, but not in the type of literals. This is because a difference in literal type may have a big impact on the refactorability of the cloned fragment. When we refactor literals that have both the same type, in case of an ``Extract Method'' refactoring, we create a method argument for this literal and pass the corresponding literal from cloned locations. However, if two literals have different types, this might not be possible (or will have a negative effect on the design of the system).

\noindent \textbf{Variable variability}\\

\subsubsection{Allow any change in some identifiers}
- Identifier of method declarations
- Identifier of locally declared variables
- Identifier of class/interface/enum declaration
- etc?

>>>>

\subsection{Type 1R clones} \label{chap:type1rclones}
We propose an alternative definition of type 1 clones. This definition requires cloned fragments to be not just textually equal, but also functionally equal. Although requiring fragements to be functionally equal, type 1R clones do not allow for change in implementation (like type 4 clones). We check functional equality of two fragments by validating the equality of the fully qualityfied identifier for referenced types, methods and variables. Type 1R clones are a subset of type 1 clones.

@@

Because of this, for type 1R clones, we compare the fully qualified method signature for all method references. A fully qualified method signature consists of the fully qualified name of the method, the fully qualified type of the method plus the fully qualified type of each of its arguments. For instance, an \texttt{eat} method could become \texttt{com.sb.AppleCore com.sb.fruitgame.Apple.eat(com.sb.fruitgame.Tool)}.

\subsubsection{Variables}
In typed programming languages, each variable declaration should declare a name and a type. When we reference a variable, we only use its name. If, in different code fragments, we use variables with the same name but different types, the code can be functionally unequal but still textually equal. As an example, see the code in figure \ref{fig:type2variables}.

\begin{figure}[H]
  \includegraphics[width=1\columnwidth]{img/type2}
  \caption{Variables with different types but the same name.}
  \label{fig:type2variables}
\end{figure}

The body of both methods in figure \ref{fig:type2variables} is equal. However, their functionality is not. The first method adds two numbers together and the other concatenates an integer to a String.

@@

\subsection{Type 2R clones}
Type 2R clones are modelled after type 2 clones, which allow any change in identifiers, literals, types, layout, and comments. For refactoring purposes, this definition is unsuitable; if we allow any change in identifiers, literals, and types, we cannot distinguish between different variables, different types and different method calls anymore. This could render two methods that have an entirely different functionality as clones (as shown in figure \ref{fig:type2} previously). Refactoring such clones can be harmful instead of helpful.

We tackle these problems with type 2R clones to be able to detect such clones that can and should be refactored. Type 1R clones are a subset of type 2R clones. All rules that apply to type 1R clones also apply to type 2R clones. Additionally, type 2R clones allow variability in literals, variables and method calls. This variability however is constrained by a threshold. Type 2R do not allow any variability in types, as opposed to type 2 clones which do allow variability in types. As type 2R clones are per definition more permissive than type 1R clones, type 1R clones are a subset of type 2R clones.

@@

\paragraph{Literal and variable variability}
We allow only variability in the value of literals and variables, but not in their types. This is because a difference in literal/variable type may have a big impact on the refactorability of the cloned fragment. When we refactor different literals/variables that have both the same type, in case of an ``Extract Method'' refactoring, we have to create a parameter for this literal and pass the corresponding literal/variable from cloned locations. However, if two literals have different types, this might not be possible (or will have a negative effect on the design of the system). This is because a lot of variability in literals will result in more parameters required in the extracted method, which is detrimental for the design of the system.

Consider the example in figure \ref{fig:type2literal1}. In this example, the two methods have two literals that differ between them. We can perform an ``Extract method'' refactoring on these to get the result in figure \ref{fig:type2literal2}. In this process, we create a method parameter for the corresponding literal.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.6\columnwidth]{img/type2literal1}
  \caption{Literal variability.}
  \label{fig:type2literal1}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.6\columnwidth]{img/type2literal2}
  \caption{Literal variability refactored.}
  \label{fig:type2literal2}
\end{figure}

\paragraph{Method call variability}

\paragraph{Allow any change in some identifiers}
- Identifier of method declarations
- Identifier of locally declared variables
- Identifier of class/interface/enum declaration
- etc?
